---
title: "Assignment 4"
author: "Ciara Stevenson"
date: "October 7, 2019"
output: html_document
---
  
  ```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE, warning = FALSE)

```{r setup, include=FALSE, message=F}
library(tidyr)
library(gapminder)
library(dplyr)
library(ggplot2)
library(readr)
```

## Exercise 1: Univariate Data Reshaping
### Univariate Option 1

#### Part 1: Life Expectancy of Scandinavian Countries (1952-2007)
```{r}
Scandinavian_LifeExp <- gapminder %>%
  filter(country == "Denmark" | country == "Sweden" | country == "Finland" | country == "Iceland") %>%
  pivot_wider(id_cols = year,
              names_from = country,
              values_from = lifeExp)
Scandinavian_LifeExp

```

#### Part 2: Scatterplot of Scandinavian Country Life Expectancies (1952-2007)
```{r}
PlotData <- reshape2::melt(Scandinavian_LifeExp, id.vars = "year") %>%
  rename(Country = variable)
ggplot(PlotData, aes(year, value, col=Country)) +
  geom_point () + 
  facet_wrap(~Country) +
  xlab("Year") +
  ylab("Life Expectancy") +
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```
#### Part 3: Relengthened Data

```{r}
Relengthened_data <- Scandinavian_LifeExp %>%
  pivot_longer(cols = c(Denmark, Finland, Iceland, Sweden),
               names_to = "country",
               values_to = "life_expectancy")
Relengthened_data
```

## Exercise 2: Multivariate Option 1

### 1. Wider Format

Make a tibble with one row per year, and columns for life expectancy and GDP per capita for China and Canada.

```{r}
gapminder2 <- gapminder %>%
  filter(country %in% c("China","Canada")) %>% 
  select(year, country, lifeExp, gdpPercap) %>% 
  pivot_wider(id_cols = year,
              names_from = country,
              values_from = c(lifeExp, gdpPercap)) 
DT::datatable(gapminder2)
```

### 2. Re-lengthen

```{r,fig.pos="center"}
gapminder2 %>%
  pivot_longer(cols = c(2:5),
               names_to = c(".value", "country"),
               names_sep = "_") %>% 
  arrange(country) %>% 
  DT::datatable()


#Exercise 3: Tibble joins

```{r, warning = FALSE, message = FALSE, echo = FALSE}
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
```

##3.1
For each guest in the guestlist (guest tibble), add a column for email address, which can be found in the email tibble.

```{r}
emailnew <- separate_rows(email, guest, sep = ", ") %>% 
  rename(name = guest) #separate guests' names and rename 'name' column
guest %>% 
  left_join(emailnew, by = "name") %>% 
  DT::datatable()
```

##3.2
Who do we have emails for, yet are not on the guestlist?

```{r}
emailnew %>% 
  anti_join(guest, by = "name") %>% 
  DT::datatable()
```

#Turner, Albert, and Vivian have emails but are not on the guestlist. 

##3.3
Make a guestlist that includes everyone we have emails for (in addition to those on the original guestlist).

```{r}
guest %>% 
  full_join(emailnew, by = "name") %>% 
  DT::datatable()
```


